---
title: "anem-vignette"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{anema-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=4, 
  fig.height=2.5
)
devtools::load_all()
```

This package was created to evaluate the hydraulic relationships among wells, in order to estimate the effect of a group of wells on drawdown at these wells or other wells. It it based on method of images from the analytical element modeling approach, in which the 2-dimensional characteristics of aquifers are reproduced by strategically placing wells within the domain.

This package models *simple* aquifer and well configurations. The boundaries of the aquifer can be specified as *no flow* or *constant head* boundaries, and the corners of the aquifer must be right angles. For fully bounded aquifers, this means that the aquifer must be a rectangle. The constant head boundaries take the head of the undisturbed aquifer, h0.

Units are designed for length dimensions given in meters (or m^2^), and time in seconds

As you will see in this vignette, the package has a number of functions that are used to set up the aquifer boundaries and wells. For practical purposes, the package was designed for simple but real aquifers, and wells and boundaries of the aquifers can be imported as shapefiles and then prepared. 

The first section, however, describes how to obtain drawdown relationships between groups of wells.

## A simple example of drawdown relationships

Load packages for the vignette

```{r setup, message=FALSE,warning=FALSE}
library(tidyverse)
library(sf)
# devtools::load_all("~/Projects/R_packages/anem")
# devtools::install_github("https://github.com/gopalpenny/anem")
# library(anem)
```

Aquifers in this package are characterized by:

* Aquifer type, confined or unconfined
* Saturated hydraulic conductivity, Ksat
* Resting hydraulic head, h0
* Thickness of the aquifer, z0 (confined aquifers only)
* Aquifer boundaries, characterized by:
  + Boundary ID, or bID
  + Coordinates in a cartesian plane, given as slope, m, and intercept, b, and coordinates (x1, y1, x2, y2)
  + Boundary type, NF or CH
* Additional user-defined parameters

Use the `define_aquifer()` function to create a new aquifer, which has the class `aquifer`. This class is essentially a list with names parameters and a separate print method.

```{r}
# define aquifer
bounds_df <- data.frame(bound_type=c("CH","NF","NF","NF"),m=c(Inf,0,Inf,0),b=c(0,1000,1000,0))
aquifer_unconfined <- define_aquifer("unconfined",1e-3,bounds=bounds_df,h0=100)
print(aquifer_unconfined)
```

Wells are characterized by:

* Well ID, or wID
* Location x, y
* Pumping rate (+ for injection, - for pumping), Q
* Radius of influence, R
* Diameter, diam
* Well images (which are noted in the `well_image` column)
* Additional user-defined columns, including groups or weights.

Define pumping wells using random locations and set a "border" between two countries at y = 500 m. This is done manually, below, by randomly generating well locations and setting wells with y > 500 as country "B". The `define_wells()` function ensures that the wells have all appropriate columns, and the class of the returned object is a `tibble` (which functions like a `data.frame`, but has a couple bells and whistles). Also note that the radius of influence (R) is defined arbitrarily, but there is also a function to calculate this manually -- see `?get_ROI` for more details. 

The `generate_image_wells()` function generates well images to recreate the bounderies defined by `aquifer_unconfined`.

```{r set_simple_aquifer}
# define wells and well images
set.seed(30)
wells_df <- data.frame(x=runif(8,0,1000),y=runif(8,0,1000),diam=1) %>%
  mutate(R=1000,  #
         country=factor(y>500,levels=c(F,T),labels=c("A","B"))) %>%
  group_by(country) %>%
  mutate(weights=1,Q=1/n()) %>% group_by()
wells <- define_wells(wells_df) %>% generate_image_wells(aquifer_unconfined)
print(wells)
```

Plot the aquifer and wells.

```{r set_simple_wells}
ggplot() +
  geom_segment(data=aquifer_unconfined$bounds,aes(x1,y1,xend=x2,yend=y2,color=bound_type)) +
  geom_abline(slope=0,intercept=500,linetype="dashed") +
  geom_point(data=wells %>% filter(wID==orig_wID),aes(x,y,fill=country),shape=21) +
  coord_equal()
```

Get drawdown relationships using `get_drawdown_relationships()`. This function calculates the average drawdown at wells in each group defined by the column `group_column`. The average is taken as the weighted mean, determined by the `weights_column`. The weights were previously set equal for all wells so that the result here is a simple mean.

```{r drawdown_relationships,results='hide'}
drawdown_relationships <- get_drawdown_relationships(wells, aquifer_unconfined, group_column=country, weights_column=weights)
```

The following results show $\Phi_{ii}$ and $\Phi_{ij}$.
```{r print_drawdown_kable,echo=FALSE}
drawdown_relationships %>% knitr::kable("html") %>% kableExtra::kable_styling()
```

## Additional examples using shapefile data inputs from Switzerland and France

## Data prep

### Examine the data

Let's begin by loading some data to practice with. The data are stored in two shapefiles, one `wells` shapefile containing *point* features with attributes for `rate` and `diameter`, and a second `boundaries` shapefile with *line* geometries and the property `bound_type`.

```{r data_import}
pts_wgs <- genevese_wells
pts_wgs

boundaries_wgs <- genevese_bounds
boundaries_wgs
```

### Plot the data

The coordinates of objects should be in a planar coordinate system in meters. Tranform both of the shapefiles to UTM projection, first by getting the UTM zone and proj4string transforming the shapefiles.

```{r transform_to_planar}
utm_zone <- longitude_to_utm_zone(mean(st_coordinates(pts_wgs)[,'X']))
utm_proj4 <- utm_zone_to_proj4(utm_zone)
pts_utm <- pts_wgs %>% sf::st_transform(utm_proj4)
boundaries_utm <- boundaries_wgs %>% sf::st_transform(utm_proj4)

ggplot() + 
  geom_sf(data=boundaries_utm, aes()) + 
  geom_sf(data=pts_utm, aes())
```

## Two wells in a uniform confined aquifer

### Prepare the aquifer 

We now need to extract the coordinates of the wells and boundaries, as well as get the slope (m) and intercept (b) of the boundaries. This can be done with the `define_aquifer()` function, which creates an `aquifer` (S3 class) object. This function also prepares boundaries using the `define_bounds()` function (e.g., run the commented code to see how it works). The `aquifer` class is essentially a list with properties specific to the aquifer, including Ksat, bounds, etc. It also has its own `print` method. See `?define_aquifer` for more details. Also note that by default, the `define_bounds` functions takes irregular quadrangles and turns them into rectangles. The method of images **only** works for aquifer with rectangular geometries (i.e., boundaries at right angles).

```{r prep_dataframes}
# bounds <- define_bounds(boundaries_utm)
aquifer_confined <- define_aquifer("confined",1e-3,bounds=boundaries_utm,h0=100,z0=10)
print(aquifer_confined)
```


### Prepare the wells 

We also need to prepare the wells. Generate an aquifer polygon and take only the wells that intersect the aquifer. 

```{r prep_wells}
bounds_polygon <- aquifer_confined$bounds %>% 
  summarize() %>% st_polygonize() %>% st_collection_extract("POLYGON")

wells_df_in_aquifer <- pts_utm %>% st_intersection(bounds_polygon) %>% group_by() 
```

Now calculate the radius of influence of each well using `get_ROI` and plot the results. The function `get_ROI` has a few different options for calculating the radius of influence of wells.

```{r define_wells}
wells_df <- wells_df_in_aquifer %>%
  mutate(Q=-1e-1/n(),
         Q=if_else(well_group=="COMMUNE",-2*Q,Q))  %>%
  mutate(R=get_ROI(Ksat=1e-4,h=100,t=3600*24*365*20,n=0.4,method="aravin-numerov")) 
wells <- define_wells(wells_df)

ggplot() + 
  geom_sf(data=boundaries_utm, aes(), linetype="dashed",alpha=0.5,color="gray") + 
  geom_sf(data=aquifer_confined$bounds, aes(color=as.factor(round(m,1)))) + 
  geom_point(data=wells, aes(x,y))
```

### Generate hydraulic head

The `wells` object contains a row for each well. Because of the principle of superposition,
the hydraulic head at any given location is the sum of the effect on hydraulic head of all pumping wells.

First define set of coordinates with which to calculate head by defining a bounding box (`grid_bounds`) and
using `expand.grid` and `seq` to generate evenly spaced points in the variable `grid_pts_head`. Then, use
these points and the `well_images` to generate hydraulic head using `get_hydraulic_head()`.

```{r hydraulic_head_confined,fig.width=6,fig.height=3.5}
grid_bounds <- aquifer_confined$bounds %>% st_coordinates() %>% as_tibble() %>%
  summarize(xmin=min(X),xmax=max(X),ymin=min(Y),ymax=max(Y))
grid_pts_head <- with(grid_bounds,crossing(x=seq(xmin,xmax,length.out=150),y=seq(ymin,ymax,length.out=100)))
heads <- get_hydraulic_head(grid_pts_head,wells,aquifer_confined) # h0=h0,Ksat=Ksat,z0=z0,aquifer_type="confined")
head_df <- grid_pts_head %>% bind_cols(head_m=heads)

ggplot(head_df) + 
  geom_raster(aes(x,y,fill=head_m))  + 
  geom_sf(data=wells, aes(size=abs(Q)),shape=16) +
  geom_contour(aes(x,y,z=head_m),bins=15,color="gray",linetype="dashed")  + 
  coord_sf(xlim=c(grid_bounds$xmin,grid_bounds$xmax),ylim=c(grid_bounds$ymin,grid_bounds$ymax)) +
  scale_size_continuous(limits=c(3,4)) +
  scale_shape_manual(values=c(1,16),limits=c(3,4))
```

### Get flow direction

Getting flow direction is similar to getting hydraulic head -- the function `get_flowdir()` is a wrapper
around `get_hydraulic_head()` that incorporates a numeric derivative from the package `numDeriv`. Similar
to before, define set of coordinates of evenly spaced points. Then, use
these points and the `well_images` to generate hydraulic head using `get_flowdir()`.

```{r flowdir_confined, message=FALSE, warning=FALSE,fig.width=6,fig.height=3.5}
grid_pts_flowdir <- with(grid_bounds,crossing(x=seq(xmin,xmax,length.out=15),y=seq(ymin,ymax,length.out=10)))
flow_dir <- get_flowdir(grid_pts_flowdir,wells,aquifer_confined)
# the dx_norm and dy_norm are used to make the arrows visible in the plot that follows.
flow_dir_df <- grid_pts_flowdir %>% bind_cols(flow_dir) %>%
  mutate(angle=atan(dy/dx),
         mag=sqrt(dx^2+dy^2),
         mag_norm=mag^(1/2)*2e4, # this scales the flow for the arrow plot
         dx_norm=cos(angle)*sign(dx)*mag_norm,
         dy_norm=sin(angle)*sign(dx)*mag_norm,
         x2=x+dx_norm,
         y2=y+dy_norm)

# plot the flow velocities
ggplot(flow_dir_df) + 
  geom_raster(aes(x,y,fill=mag_norm))  + 
  geom_sf(data=wells, aes(),size=3) +
  coord_sf(xlim=c(grid_bounds$xmin,grid_bounds$xmax),ylim=c(grid_bounds$ymin,grid_bounds$ymax))
```

Now, plot the results together. 
```{r plot_confined,fig.width=8,fig.height=4.5}
ggplot(head_df) +
  geom_raster(aes(x,y,fill=head_m))  +
  geom_point(data=wells, aes(x,y),shape=1,stroke=1) +
  geom_contour(aes(x,y,z=head_m),bins=20,color="gray",linetype="dashed")  + 
  geom_segment(data=flow_dir_df,
               aes(x,y,xend=x2,yend=y2),
               arrow = arrow(ends="last",type="closed",length=unit(1,"mm"))) +
  coord_equal() +
  # coord_sf(xlim=c(grid_bounds$xmin,grid_bounds$xmax),ylim=c(grid_bounds$ymin,grid_bounds$ymax),crs = st_crs(wells)) +
  scale_shape_manual(values=c(1,16),limits=c(3,4)) #+ scale_size_continuous(limits=c(3,4))
```

## Two wells in an aquifer with bounds on all sides

Here we will use the `wells` object generated for the confined aquifer. However, we need to define a new unconfined auifer. This is done again using the `define_aquifer` function.

```{r prep_unconfined}
aquifer_unconfined <- define_aquifer("unconfined",Ksat=0.0001,h0=100,bounds=boundaries_utm)
```

### Generate well images

The `aquifer_unconfined` has boundaries defined, but these boundaries won't actually reproduce the behavior of boundaries in the groundwater model. To reproduce the behavior of the boundaries, we need to generate image wells by mirroring wells across each of the boundaries. Wells are mirrored across no-flow ("NF") boundaries with the inverse rate of pumping, and across constant head ("CH") boundaries with the same pumping rate.

* This mirroring process only works with rectangular boundaries.
  + Use only straight lines.
  + No complex shapes.
  
Lastly, to know the effect of a well on any given location, we need to know the radius of influence of the well. This package provides a couple ways to do this through the `get_ROI` function. See `?get_ROI` for more resources. 

```{r well_imaging,fig.width=6,fig.height=3.5}
well_images <- generate_image_wells(wells,aquifer_unconfined) 
well_roi <- gen_circles(well_images %>% rename(r=R))

ggplot()  +
  geom_sf(data=aquifer_unconfined$bounds, aes()) + 
  geom_sf(data=boundaries_utm, aes(), linetype="dashed",alpha=0.5,color="gray") + 
  geom_point(data=well_images,aes(x,y),alpha=0.5) + 
  # geom_polygon(data=well_roi,aes(x,y,group=id),alpha=0.3,color="black")  
  geom_polygon(data=well_roi,aes(x,y,group=id),alpha=0.01)

```

Some of the images wells are too far away to affect the aquifer. Select only the wells that have a mirror distance less than the radius of influence of the wells.

```{r well_imaging_filter,fig.width=6,fig.height=3.5}
# well_images <- well_images %>% filter(max_mirror_dist<R)
# well_roi <- gen_circles(well_images %>% rename(r=R))
ggplot()  +
  geom_sf(data=boundaries_utm, aes(), linetype="dashed",alpha=0.5,color="gray")+
  geom_polygon(data=well_roi,aes(x,y,group=id),alpha=0.1,color="black") +
  geom_sf(data=aquifer_unconfined$bounds, aes(color=as.factor(bID)),size=1) +
  geom_point(data=well_images,aes(x,y,fill=Q),alpha=0.5,size=3,shape=21) +
  scale_fill_gradient2("Pumping rate",low="blue",mid="gray",high="red") +
  scale_color_discrete("Boundary")
```

### Check to ensure the boundaries are working as intended

The function `get_bounds_behavior()` is a helper function to generate hydraulic properties at the boundaries. It obtains hydraulic head and the flow normal to the boundary. Normal flow is defined such that positive flow has some component in the x-direction (the y-direction depends on the normal vector for the boundary). The function `plot_bounds_behavior()` is a wrapper around `get_bounds_behavior()`, and it can be used to quickly compare the hydraulic head along the boundaries and flow across the boundaries under scenarios: 

* no images (ie, homogeneous, uniform aquifer -- all well images removed)
* images (ie, aquifer with boundaries -- all well images intact)

```{r check_bounds,fig.width=6,fig.height=4}
gg_bounds_behavior <- plot_bounds_behavior(well_images,aquifer_unconfined)
gridExtra::grid.arrange(gg_bounds_behavior$p_h,gg_bounds_behavior$p_f,nrow=1)
```

### Generate hydraulic head

The `well_images` object contains a row for each well (including actual wells and well images). Because of the principle of superposition, the hydraulic head at any given location is the sum of the effect on hydraulic head of all pumping wells.

First define set of coordinates with which to calculate head by defining a bounding box (`grid_bounds`) and
using `expand.grid` and `seq` to generate evenly spaced points in the variable `grid_pts_head`. Then, use
thes points and the `well_images` to generate hydraulic head using `get_hydraulic_head()`.

```{r hydraulic_head_unconfined,fig.width=7,fig.height=3}
grid_bounds_unconf <- aquifer_unconfined$bounds %>% st_coordinates() %>% as_tibble() %>%
  summarize(xmin=min(X),xmax=max(X),ymin=min(Y),ymax=max(Y))
# grid_bounds_unconf <- data.frame(xmin=265000,xmax=290000,ymin=5108000,ymax=5120000)
grid_pts_head_unconf <- with(grid_bounds_unconf,crossing(x=seq(xmin,xmax,length.out=150),y=seq(ymin,ymax,length.out=100)))
heads_unconf <- get_hydraulic_head(grid_pts_head_unconf,well_images,aquifer_unconfined)
head_unconf_df <- grid_pts_head_unconf %>% bind_cols(head_m=heads_unconf)

ggplot(head_unconf_df %>% mutate(influenced_by_wells=head_m!=100)) + 
  geom_raster(aes(x,y,fill=influenced_by_wells)) + 
  geom_sf(data=aquifer_unconfined$bounds, aes()) + 
  ggtitle("Region of well influence")

ggplot(head_unconf_df) + 
  geom_raster(aes(x,y,fill=head_m))  + 
  geom_contour(aes(x,y,z=head_m),bins=20,color="gray",linetype="dashed")  + 
  geom_sf(data=aquifer_unconfined$bounds, aes()) + 
  # geom_sf(data=boundaries_utm, aes(), linetype="dashed",alpha=0.5,color="gray") + 
  geom_point(data=well_images %>% filter(well_image=="Actual"),aes(x,y)) +
  # geom_sf(data=pts_utm, aes(size=abs(Q),shape=type)) +
  coord_sf(xlim=c(grid_bounds_unconf$xmin,grid_bounds_unconf$xmax),ylim=c(grid_bounds_unconf$ymin,grid_bounds_unconf$ymax)) +
  scale_shape_manual(values=c(1,16),limits=c(3,4))
```

### Get flow direction

Getting flow direction is similar to getting hydraulic head -- the function `get_flowdir()` is a wrapper
around `get_hydraulic_head()` that incorporates a numeric derivative from the package `numDeriv`. Similar
to before, define set of coordinates of evenly spaced points. Then, use
these points and the `well_images` to generate hydraulic head using `get_flowdir()`.

```{r flowdir_unconfined, message=FALSE, warning=FALSE,fig.width=6,fig.height=3.5}
grid_pts_flowdir_unconf <- with(grid_bounds_unconf,crossing(x=seq(xmin,xmax,length.out=30),y=seq(ymin,ymax,length.out=20)))
flow_dir_unconf <- get_flowdir(grid_pts_flowdir_unconf,well_images,aquifer_unconfined)
# the dx_norm and dy_norm are used to make the arrows visible in the plot that follows.
flow_dir_unconf_df <- grid_pts_flowdir_unconf %>% bind_cols(flow_dir_unconf) %>%
  mutate(angle=atan(dy/dx),
         mag=sqrt(dx^2+dy^2),
         mag_norm=mag^(1/3)*5e3,
         dx_norm=cos(angle)*sign(dx)*mag_norm,
         dy_norm=sin(angle)*sign(dx)*mag_norm,
         x2=x+dx_norm,
         y2=y+dy_norm)
# flow_dir_check <- flow_dir_df %>% select(dx,dx_norm,dy,dy_norm) %>%
#   mutate(check=sign(dx)==sign(dx_norm) & sign(dy)== sign(dy_norm))

# plot the flow velocities
ggplot(flow_dir_unconf_df) + 
  geom_raster(aes(x,y,fill=mag_norm))  + 
  geom_sf(data=aquifer_unconfined$bounds, aes()) + 
  geom_sf(data=boundaries_utm, aes(), linetype="dashed",alpha=0.5,color="gray") + 
  geom_point(data=well_images %>% filter(well_image=="Actual"),aes(x,y)) +
  coord_sf(xlim=c(grid_bounds_unconf$xmin,grid_bounds_unconf$xmax),
           ylim=c(grid_bounds_unconf$ymin,grid_bounds_unconf$ymax)) +
  scale_shape_manual(values=c(1,16),limits=c(3,4))
```


Now, plot the results together. 
```{r plot_unconfined,fig.width=6,fig.height=3.5}
ggplot(head_unconf_df) +
  geom_raster(aes(x,y,fill=head_m))  +
  geom_sf(data=aquifer_unconfined$bounds, aes()) + 
  geom_contour(aes(x,y,z=head_m),bins=20,color="gray",linetype="dashed")  + 
  # geom_sf(data=boundaries_utm, aes(), linetype="dashed",alpha=0.5,color="gray") + 
  geom_point(data=well_images %>% filter(well_image=="Actual"),aes(x,y)) +
  geom_segment(data=flow_dir_unconf_df,
               aes(x,y,xend=x2,yend=y2),arrow = arrow(ends="last",type="closed",length=unit(1,"mm"))) +
  coord_sf(xlim=c(grid_bounds$xmin,grid_bounds$xmax),ylim=c(grid_bounds$ymin,grid_bounds$ymax)) +
  scale_shape_manual(values=c(1,16),limits=c(3,4)) #+ scale_size_continuous(limits=c(3,4))
```

## Calculate Dii, Dij, and ri.

I will write a function to do this in the future, but it's not too difficult to get started with this.

Set the total pumping rate for each group equal to 1

```{r generate_game_wells}
wells_genevese <- wells %>% filter(well_image=="Actual")
aquifer_genevese <- aquifer_unconfined
wells_game <- wells_genevese %>%
  mutate(player=factor(well_group,levels=c("SIG","France","Commune"),labels=c("S","F","Recharge")),
         well_sign=if_else(well_type=="pumping",-1,1)) %>%
  group_by(player) %>% 
  mutate(weights=1,Q=1/n()) %>% group_by()
well_game_images <- generate_image_wells(wells_game,aquifer_genevese)
```

Calculate Dij as the average effect of wells j on wells i. Note that wells j must include all imaged wells, while the effect is only on the actual wells.

```{r well_relationships,fig.width=6,fig.height=3.5}
drawdown_genevese <- get_drawdown_relationships(well_game_images,aquifer_genevese,player,weights)
```

```{r print_drawdown_genevese,echo=FALSE}
drawdown_genevese %>% knitr::kable() %>% kableExtra::kable_styling()
```


```{r}
ggplot() +
  geom_sf(data=aquifer_unconfined$bounds, aes()) + 
  geom_point(data=wells_game,aes(x,y,color=player,shape=well_type),size=3,stroke=1) +
  scale_shape_manual(values=c(1,16)) #+ scale_size_continuous(limits=c(3,4))
```

