---
title: "anem-demonstration"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{anem-demonstration}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=4, 
  fig.height=2.5
)
```

```{r}
library(anem)
```


### Abstract 

Numerous tools exist for simulating groundwater flow via finite element models (e.g., MODFLOW) or analytic element models (e.g., gflow).
These tools allows for analysis of complex groundwater behavior, functionality that may not be necessary projects that demand simple groundwater models and more advanced statistical or theoretical analysis.
Tighter integration of groundwater models with analytical environments can provide numerous benefits for quickly analyzing groundwater hydrology problems.
To this end, we develop an R package designed to simulate 2-dimensional, steady-state groundwater flow in confined and unconfined aquifers with rectangular boundaries using analytic elements and the method of well images.
The package, "anem", contains functions for defining aquifers and pumping wells, determining hydraulic head and lateral flow, and calculating drawdown relationships among groups of wells.
These features make this package a useful tool for simulating groundwater behavior as a precursor to more complex groundwater modeling, or simulating synthetic scenarios as a way to explore variability and uncertainty.
Because R is a functional programming language, the outputs from this package can be easily analyzed, mapped, and incorporated into more complex models in R.
Indeed, the linear behavior of groundwater flow and principle of superposition means that physically plausible groundwater dynamics can be incorporated into more complex models of hydrology and coupled human-natural systems.
We validate the package and describe multiple scenarios in which this package might be used, emphasizing the potential benefits of incorporating drawdown relationships into more complex models.

### Method of images

The package uses the method of images, where wells are mirrored across aquifer boundaries to reproduce the characterics of the boundary, no flow or constant head:

```{r echo=FALSE, message=FALSE,warning=FALSE}
library(tidyverse)
library(sf)
# devtools::install_github("https://github.com/gopalpenny/anem")
# devtools::load_all("~/Projects/R_packages/anem")
library(anem)
```

```{r echo=FALSE, message=FALSE,warning=FALSE,fig.width=5.5,fig.height=3}
# Create a grid of locations
loc <- crossing(x=seq(-200,200,length.out=201),y=seq(-200,200,length.out=201))

# Constant head boundary
wells_constant_head <- define_wells(x=c(-100,100),y=c(-0,0),Q=c(1e-2,-1e-2),diam=c(0.1,0.1),R=c(500,500))
aquifer <- define_aquifer("confined",1e-4,z0=20,h0=0)
constant_head_boundary <- loc %>%
  dplyr::bind_cols(streamfunction=get_stream_function(loc,wells_constant_head,aquifer)) %>%
  dplyr::bind_cols(head=get_hydraulic_head(loc,wells_constant_head,aquifer))

# No flow boundary
wells_no_flow <- define_wells(x=c(-100,100),y=c(-0,0),Q=c(-1e-2,-1e-2),diam=c(0.1,0.1),R=c(500,500))
no_flow_boundary <- loc %>%
  dplyr::bind_cols(streamfunction=get_stream_function(loc,wells_no_flow,aquifer)) %>%
  dplyr::bind_cols(head=get_hydraulic_head(loc,wells_no_flow,aquifer))

thm <- theme(axis.text=element_blank(),axis.title=element_blank(),axis.ticks=element_blank(),
             panel.background = element_rect(fill=NA,color="black"),
             panel.border = element_rect(fill=NA,color="black"),
             plot.title = element_text(size=10))

p_CH <- ggplot() +
  geom_contour(data=constant_head_boundary,aes(x,y,z=head),bins=50,linetype="dashed",color="black") +
  geom_contour(data=constant_head_boundary,aes(x,y,z=streamfunction),bins=25,color="darkblue") +
  geom_point(data=wells_constant_head,aes(x,y,fill=well_type),size=3,shape=21) +
  theme(legend.position=c(0.8,0.15),legend.title=element_blank()) +
  coord_equal() + thm + ggtitle("Constant head boundary at x = 0")
p_NF <- ggplot() +
  geom_contour(data=no_flow_boundary,aes(x,y,z=head),bins=25,linetype="dashed",color="black") +
  geom_contour(data=no_flow_boundary,aes(x,y,z=streamfunction),bins=50,color="darkblue") +
  geom_point(data=wells_no_flow,aes(x,y,fill=well_type),size=3,shape=21) +
  theme(legend.position=c(0.8,0.1),legend.title=element_blank()) +
  coord_equal() + thm + ggtitle("No flow boundary at x = 0")
gridExtra::grid.arrange(p_CH,p_NF,nrow=1)
```


### Drawdown relationships

```{r setup, include=FALSE}
library(tidyverse)
library(sf)
# devtools::install_github("https://github.com/gopalpenny/anem")
# devtools::load_all("~/Projects/R_packages/anem")
library(anem)
```

Use the `define_aquifer()` function to create a new aquifer, which has the class `aquifer`. This class is essentially a list with names parameters and a separate print method.

```{r}
# define aquifer
aquifer <- define_aquifer("unconfined",1e-3,bounds=bounds_example,h0=100)
```

The `generate_image_wells()` function generates well images to recreate the bounderies defined by `aquifer`. The `wells_example` includes 8 wells, with 4 each associated with country "A" and "B", respectively.

```{r set_simple_aquifer}
# define wells and well images
wells_actual <- define_wells(wells_example) 
wells <- wells_actual %>% generate_image_wells(aquifer)
```

Plot the aquifer and wells.

```{r set_simple_wells,echo=FALSE}
ggplot() +
  geom_segment(data=aquifer$bounds,aes(x1,y1,xend=x2,yend=y2,color=bound_type)) +
  geom_abline(slope=0,intercept=500,linetype="dashed") +
  geom_point(data=wells_actual,aes(x,y,fill=country),shape=21) +
  coord_equal()
```

Get drawdown relationships using `get_drawdown_relationships()`. This function calculates the average drawdown at wells in each group defined by the column `group_column`. The average is taken as the weighted mean, determined by the `weights_column`. The weights were previously set equal for all wells so that the result here is a simple mean. The results show $\Phi_{ii}$ and $\Phi_{ij}$.

```{r drawdown_relationships,results='hide'}
drawdown_relationships <- get_drawdown_relationships(wells, aquifer, country, weights)
```

```{r print_drawdown_kable,echo=FALSE}
drawdown_relationships %>% 
  knitr::kable("html") #%>% kableExtra::kable_styling()
```

### Plot the head and flow

The hydrodynamics of the aquifer can be mapped by obtaining gridded heand flow using the `get_gridded_hydrodynamics()` function. The function takes as input the wells, aquifer, and grid dimensions for head and flow. It returns a `list` object with data.frames for head and flow, which can then be plotted.

```{r}
gridded <- get_gridded_hydrodynamics(wells,aquifer,c(20,20),c(8,8))
```

```{r echo=FALSE, warning=FALSE}
ggplot() +
  geom_raster(data=gridded$head,aes(x,y,fill=head_m)) +
  geom_segment(data=gridded$flow,aes(x,y,xend=x2,yend=y2),
               arrow = arrow(ends="last",type="closed",length=unit(1,"mm")),color="black") +
  geom_segment(data=aquifer$bounds,aes(x1,y1,xend=x2,yend=y2,color=bound_type)) +
  geom_point(data=wells %>% filter(wID==orig_wID),aes(x,y),shape=21) +
  coord_equal()
```

